---
output: pdf_document
---
### Peter Lee
### STATS 202
### 9/30/2015

### Homework 1, Problem 6



## Part A

#### Load the training_target.csv file into a data frame in R using the read.csv command
```{r}
tbl.target = read.csv("/Users/PeterLee/Documents/SCHOOL/Stanford/STATS 202/Homework/Homework 1/training_target.csv")
```

#### Question: How many patients are in this data frame?
```{r}
dim(tbl.target)
ncol(tbl.target)
nrow(tbl.target)
```
#### Solution: 2424 rows with 2 columns. Thus, 2424 patients.

#### Use the handy command summary to display distributional information about the training target slopes.
```{r}
summary(tbl.target)
```



## Part B

#### Now, load the training_features.csv file into a data frame in R using the read.csv command
```{r}
tbl.features = read.csv("/Users/PeterLee/Documents/SCHOOL/Stanford/STATS 202/Homework/Homework 1/training_features.csv")
```

#### Question: How many feature columns are in this data frame?
```{r}
ncol(tbl.features)
```
#### Solution: There are 858 feature columns.

#### Write code to find the number of NA values in each column, and plot a histogram of these NA counts.
```{r}
num.nas <- function(tbl.features) sum(is.na(tbl.features))
num.features.nas <- apply(tbl.features, 2, num.nas)
```

#### Solution: Can use this code to find number of NA values in each column
```{r, eval=FALSE}
num.features.nas
```

#### Solution: Histogram
```{r}
hist(num.features.nas,
     main = "Frequency of NAs in Column",
     xlab = "Counts of NAs in Column",
     xlim = c(0,2500),
     ylim = c(0,400),
     breaks = 20,
     col = "yellow",
     border = "blue")
```
     

## Part C

#### Create dummy variable for weight.slope (help from website)
```{r}
feature.name <- "weight.slope"
dummy <- paste0("is.na", feature.name)
is.na.feature <- is.na(tbl.features[,feature.name])

tbl.features[,dummy] <- as.integer(is.na.feature)

tbl.features[is.na.feature,feature.name] <- median(tbl.features[,feature.name], na.rm = TRUE)
```

#### Question: Write code that: (a) Add binary dummy variable to training features data frame whether or not variable value was NA, and (b) Replace all missing values for that variable with the median non-missing value for that variable

```{r}
feature.names <- names(tbl.features)
for (feature.name in feature.names[-1]) 
  {
dummy <- paste0("is.na.", feature.name)
is.na.feature <- is.na(tbl.features[,feature.name])

tbl.features[,dummy] <- as.integer(is.na.feature)

tbl.features[is.na.feature,feature.name] <- median(tbl.features[,feature.name], na.rm = TRUE)

  }
```

#### Solution: To demonstrate the dummy variables, the number of columns has doubled
```{r}
dim(tbl.features)
```


## Part D

#### Import last 2 CSVs
```{r}

tbl.valid.features = read.csv("/Users/PeterLee/Documents/SCHOOL/Stanford/STATS 202/Homework/Homework 1/validation_features.csv")

tbl.valid.target = read.csv("/Users/PeterLee/Documents/SCHOOL/Stanford/STATS 202/Homework/Homework 1/validation_target.csv")
```

#### Question: How many validation patients and feature columns are there?
```{r}
dim(tbl.valid.features) 
```
#### Solution: There are 858 feature columns.


#### Question: How does the summary of the validation target slopes compare with that of the training target slopes?
```{r}
summary(tbl.valid.target)
summary(tbl.target)
```
#### Solution: Summary of validation target slopes is lower on average (-0.79 vs.-0.73) than that of training target slopes.


#### Question: Display a summary of the weight.slope column of the validation features
```{r}
summary(tbl.valid.features$weight.slope)
```

                






